Analytical choices for analyzing multidimensional behavior - Many analysts test hypotheses about human speech.

### 2021-03-03 RR_manuscript/Draft_RR.Rmd
:memo: - Draft_RR edits in Introduction, plus typo fixes

In a collaborative effort, @silberzahn2018many let twenty-nine independent analysis teams address the same research hypothesis.
Analytical approaches and consequently the results varied widely between teams.
Sixty-nine percent of the teams found support for the hypothesis, and 31% did not.
Out of the 29 analytical strategies, there were 21 unique combinations of covariates.
Importantly, the observed variability was neither predicted by the team's preconceptions about the phenomenon under investigation nor by peer ratings of the quality of their analyses.
The authors results suggest that analytic plurality is a fact of life and not driven by different levels of expertise or bias.
Similar crowd-sourced studies recruiting independent analyst teams showed similar results.

<!-- TODO: SUMUP neuroscience, cognitive modeling, clinical, predictive models -->

@@

Speech is a complex signal that is characterized by structurally different acoustic landmarks distributed throughout different temporal domains.
Thus, choosing how to measure a phenomenon of interest is an important and non-trivial analytical decision. Take for example the following sentence in 1:

<!-- TR: I'd suggest an easier example (Sheila and Pat are both uncommon names, internationally speaking) and possibly a more relateable (possibly timely) example. Something about zoom meetings or sth maybe? Examples are super important to relate to the reader and lure them in. Here is my suggestion-->

@@

<!-- Now, imagine we are interested in measuring the difference between that sentence and the sentence "Sheila says *Mat* is clever".
How can we obtain a measure of dissimilarity between the first sound in *Pat* and the first sound in *Mat*? -->

<!-- TR: Keep in mind that the research question is not about segmental phenomena. To relate to the question later, I'd suggest a prosodic measure, e.g.  -->

@@

<!-- TR: I changed features to parameters. Features is too close to phonological theory and will confuse some linguists, I believe. -->

@@

(2a) "I can't bear ANOTHER meeting on zoom."
(2b) "I can't bear another meeting on ZOOM."

In text ref to Figure \@ref(fig:forkingPaths)

@@

<!-- TR: Insertion does not work yet, gotta dash. Also the following paragraph is just a draft for now -->

Do we only compare the word *another* in 2a and 2b or also the word *zoom* or do we measure utterance wide acoustic profiles? Do we measure the whole word? Or just the stressed syllable?
Do we average the domain or do we measure a specific point in time?
Do we measure fundamental frequency or intensity?
When looking at phrase-wide temporal domains, the number of possible analytical pipelines quickly explodes.
This plurality of analytical paths is illustrated in figure X.
When comparing two utterance such as 2a and 2b, there are many things to consider.
Even if we know that we want to compare fundamental frequency of only the word *another* across utterances 2a and 2b, there are still many decisions to be made, all of which can be justified.
For example, we could measure f0 at specific points in time like the onset of the window, the offset, the midpoint.
We could also measure the value or time of the minimum or maximum f0 value.
We could summarise f0 across the entire window and extract the mean, median or standard deviation of f0.
And the garden of forking paths does not stop here.
In Figure X, we went with a specific option to automatically calculate f0, INSERT SOME EXAMPLES OF PITCH TRACKING OPTIONS.
Moreover, knowing that theses estimates are somewhat noisy, we could smooth these contours to different degrees, automatically or manually remove estimates that are off, etc.

These decisions are usually made prior to any statistical analysis, but are at times revised a posteriori (i.e. after data collection and/or preliminary analyses) in light of unforeseen or surprising outcomes.
These myriads of possible decisions are exponentiated by researcher degrees of freedom related to statistical analysis [e.g. wicherts2016].
Even the analysis of a single measure can be approached via an ever-increasing range of different statistical models (REFs).
The present paper probes this garden of forking paths in the analysis of speech.
To assess the variability in data analysis pipelines across independent researchers, we provided XX analytical teams with an experimentally elicited speech corpus and asked them to investigate acoustic differences related to a functional contrast that might be manifested across the whole utterance.

@@

<!-- TODO: INFORMATION ABOUT THE DATA SET AND EXP DESIGN + TRANSITION -->
To answer the question of whether speakers modify speech to signal atypical referents, thirty native German speakers were recorded in a speech production study while interacting with a confederate (one of the experimenters).
The participants had to verbally instruct the confederate to select a specified target object out of four object presented on a screen.
The subject and confederate were seated at the opposite sides of a table, each facing one of two mirrored computer screens.
The participant and the experimenter could not see each other nor each others' screens.
After a familiarisation phase (see XXX), the subject first saw four colored objects in the top left, top right, bottom left, and bottom right corners of the screen.
One of the objects served as the target, another as the competitor, and the remaining two objects served as distractors.
Objects were referred to using noun phrases consisting of an adjective modifier denoting color and a modified object (e.g. *Gelbe Zitrone* 'yellow lemon', *Rote Gurke* 'red cucumber', *Rote Socken* 'red socks').

In the center of the screen, a black cube was displayed, which could be moved by the experimenter.
The participant would read a sentence prompt out loud (*Du sollst den Würfel auf der <COLOR OBJECT> ablegen* 'You have to put the cube on top of the <COLOR OBJECT>') to instruct the experimenter to drag the cube on top of one of the four depicted objects (the *competitor*) using the mouse.
After the experimenter had moved the cube as instructed, the subject would read another sentence prompt (*Und jetzt sollst du den Würfel auf der <COLOR OBJECT> ablegen* 'And now, you have to put the cube on top of the <COLOR OBJECT>') instructing the experimenter to move the cube on top of a different object (the *target*).

The two sentence prompts were used to create a focus contrast between the competitor and the target object.
If the competitor and target objects differed but their color did not (e.g. *yellow banana* vs *yellow tomato*), the noun was in focus (the Noun Focus condition, NF).
If the the objects were the same object but differed in color (e.g. *yellow banana* vs *blue banana*), the color adjective was discourse-pragmatically focused (the Adjective Focus condition, AF).
If both the color and the object differed (e.g. *yellow banana* vs *blue tomato*), the whole noun phrase was in focus (the Adjective/Noun Focus condition, ANF).
The NF condition constituted the experimentally relevant condition, while the AF and ANF conditions acted as fillers.
The color-object combinations in the Noun Focus (NF) condition were manipulated with respect to their typicality.
The combinations were either typical (e.g. *Orangene Mandarine* 'orange mandarin'), medium typical (e.g. *Grüne Tomate* 'green tomato'), or atypical (e.g. *Gelbe Kirsche* 'yellow cherry').
The typicality rating of each color-object combination was established, prior to the production experiment, with a norming study (see Appendix XXX for details).
Each participant went through 15 NF trials, 10 AF and 10 ANF trials.
Each trial was repeated twice, yielding a total of 70 trials ($(15 \times 10 \times 10) \times 2$) per participant and a grand-total of 2100 observations ($70 \times 30$ participants), of which 900 ($15 \times 2 \times 30$ participants) were the experimentally relevant trials.

@@

The present project examines the extent to which subjective choices by different researchers analyzing a complex speech data set affect the reported results.
We are further interested in which factors affect the researchers' final results.

@@

Furthermore, the reviewer will be given the option to include further comments in a text box.

After submitting the review, a methods section from a second analysis will then be made available to the reviewer.
This same sequence will be followed until all analyses allocated to a given reviewer have been provided and reviewed.
After providing the final review, the reviewer will be simultaneously presented with all four (or more) methods sections that the reviewer has just completed reviewing, the option to revise their original ratings, and a text box to provide for an explanation.
<!-- The invitation to revise the original ratings will be as follows: "If, now that you have seen all the analyses you are reviewing, you wish to revise your ratings of any of these analyses, you may do so now."
The text box will be prefaced with this prompt: "Please explain your choice to revise (or not to revise) your ratings." -->

## Step 4: Evaluate Variation

>>>>

In a collaborative effort, @silberzahn2018many let twenty-nine independent analysis teams address the same research hypothesis.
Analytical approaches and consequently the results varied widely between teams.
Sixty-nine percent of the teams found support for the hypothesis, and 31% did not.
Out of the 29 analytical strategies, there were 21 unique combinations of covariates.
Importantly, the observed variability was neither predicted by the team's preconceptions about the phenomenon under investigation nor by peer ratings of the quality of their analyses.
The authors results suggest that analytic plurality is a fact of life and not driven by different levels of expertise or bias.

Similar crowd-sourced studies recruiting independent analyst teams showed similar results.
Both @dutilh2019 ad @starns2019 investigated analysts choices when inferring theoretical constructs based on the same data set using computational models.
<!-- TODO: Parker et al. bib info -->
Both studies reveal vastly different modeling strategies, even though scientific conclusions were similar across analysis teams [see also Parker et al. (accepted) on analytical flexibility in ecology and @botvinik-nezer2020 on neuroimaging data].
@bastiaansen2020 crowdsourced clinical recommendations based on data analysis of an individual patient.
Results suggests that analysis team differed substantially regarding decisions related to both the statistical analysis of the data set and the theoretical rationale behind interpreting the statistical results, indicating that there are multiple layers for degrees of freedom that interact with each other.

@@

Speech is a complex signal that is characterized by structurally different acoustic landmarks distributed throughout different temporal domains.
Thus, choosing how to measure a phenomenon of interest is an important and non-trivial analytical decision. Take for example the following sentence in 1.

@@

@@

@@

(2a) "I can't bear ANOTHER meeting on zoom."

(2b) "I can't bear another meeting on ZOOM."

@@

Do we only compare the word *another* in 2a and 2b or also the word *zoom* or do we measure utterance wide acoustic profiles? Do we measure the whole word? Or just the stressed syllable?
Do we average the domain or do we measure a specific point in time?
Do we measure fundamental frequency or intensity?
When looking at phrase-wide temporal domains, the number of possible analytical pipelines quickly explodes.
This plurality of analytical paths is illustrated in figure X.
When comparing two utterance such as 2a and 2b, there are many things to consider.
Even if we know that we want to compare fundamental frequency of only the word *another* across utterances 2a and 2b, there are still many decisions to be made, all of which can be justified.
For example, we could measure f0 at specific points in time like the onset of the window, the offset, the midpoint.
We could also measure the value or time of the minimum or maximum f0 value.
We could summarise f0 across the entire window and extract the mean, median or standard deviation of f0.
And the garden of forking paths does not stop here.
<!-- TODO: add examples of pitch tracking options -->
In \highlight[id=SC, comment = {The figure text is quite small. Probably better to change to a longer rather then wider layout.}]{Figure} \@ref(fig:forkingPaths), we went with a specific option to automatically calculate f0, INSERT SOME EXAMPLES OF PITCH TRACKING OPTIONS.
Moreover, knowing that theses estimates are somewhat noisy, we could smooth these contours to different degrees, automatically or manually remove estimates that are off, etc.

These decisions are usually made prior to any statistical analysis, but are at times revised a posteriori (i.e. after data collection and/or preliminary analyses) in light of unforeseen or surprising outcomes.
These myriads of possible decisions are exponentiated by researcher degrees of freedom related to statistical analysis [e.g. wicherts2016].
<!-- TODO: references to range of statistical models -->
Even the analysis of a single measure can be approached via an ever-increasing range of different statistical models (REFs).
The present paper probes this garden of forking paths in the analysis of speech.
To assess the variability in data analysis pipelines across independent researchers, we provided XX analytical teams with an experimentally elicited speech corpus and asked them to investigate acoustic differences related to a functional contrast that might be manifested across the whole utterance.

@@

To answer the question of whether speakers modify speech to signal atypical referents, thirty native German speakers were recorded in a speech production study while interacting with a confederate (one of the experimenters).
The participants had to verbally instruct the confederate to select a specified target object out of four object presented on a screen.
The subject and confederate were seated at the opposite sides of a table, each facing one of two mirrored computer screens.
The participant and the experimenter could not see each other nor each others' screens.
After a familiarisation phase (see XXX), the subject first saw four colored objects in the top left, top right, bottom left, and bottom right corners of the screen.
One of the objects served as the target, another as the competitor, and the remaining two objects served as distractors.
Objects were referred to using noun phrases consisting of an adjective modifier denoting color and a modified object (e.g. *Gelbe Zitrone* 'yellow lemon', *Rote Gurke* 'red cucumber', *Rote Socken* 'red socks').

In the center of the screen, a black cube was displayed, which could be moved by the experimenter.
The participant would read a sentence prompt out loud (*Du sollst den Würfel auf der <COLOR OBJECT> ablegen* 'You have to put the cube on top of the <COLOR OBJECT>') to instruct the experimenter to drag the cube on top of one of the four depicted objects (the *competitor*) using the mouse.
After the experimenter had moved the cube as instructed, the subject would read another sentence prompt (*Und jetzt sollst du den Würfel auf der <COLOR OBJECT> ablegen* 'And now, you have to put the cube on top of the <COLOR OBJECT>') instructing the experimenter to move the cube on top of a different object (the *target*).
The audio of this second utterance in the trial was recorded.

The two sentence prompts were used to create a focus contrast between the competitor and the target object.
If the competitor and target objects differed but their color did not (e.g. *yellow banana* vs *yellow tomato*), the noun was in focus (the Noun Focus condition, NF).
If the the objects were the same object but differed in color (e.g. *yellow banana* vs *blue banana*), the color adjective was discourse-pragmatically focused (the Adjective Focus condition, AF).
If both the color and the object differed (e.g. *yellow banana* vs *blue tomato*), the whole noun phrase was in focus (the Adjective/Noun Focus condition, ANF).
The NF condition constituted the experimentally relevant condition, while the AF and ANF conditions acted as fillers.
The color-object combinations in the Noun Focus (NF) condition were manipulated with respect to their typicality.
The combinations were either typical (e.g. *Orangene Mandarine* 'orange mandarin'), medium typical (e.g. *Grüne Tomate* 'green tomato'), or atypical (e.g. *Gelbe Kirsche* 'yellow cherry').
The typicality rating of each color-object combination was established, prior to the production experiment, with a norming study (see Appendix XXX for details).

Each participant went through 15 NF trials, 10 AF and 10 ANF trials.
Each trial was repeated twice, yielding a total of 70 trials ($(15 \times 10 \times 10) \times 2$) per participant and a grand-total of 2100 observations ($70 \times 30$ participants), of which 900 ($15 \times 2 \times 30$ participants) were the experimentally relevant trials.
The recordings of the second sentence prompt of each trial make up the speech dataset given to the analysis teams.

@@

The present study examines the extent to which subjective choices by different researchers analyzing a complex speech data set affect the reported results.
We are further interested in which factors affect the researchers' final results.

@@

Furthermore, the reviewers will be given the option to include further comments in a text box.

After submitting the review, a methods section from a second analysis will then be made available to the reviewer.
This same sequence will be followed until all analyses allocated to a given reviewer have been provided and reviewed.
After providing the final review, the reviewer will be simultaneously presented with all four (or more) methods sections that the reviewer has just completed reviewing, the option to revise their original ratings, and a text box to provide for an explanation.

## Step 4: Evaluate variation
